Grove: A Convergent Collaborative Structure-Editor Calculus

### 2020-10-29 grove-paper.tex
Daily revision (Michael D. Adams)

\tikzset{
  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
}
\tikzset{edge style/.style={
  draw,
  font={\small\ttfamily},
  /tikz/every edge quotes/.style={
    %draw=gray!20,
    anchor=west,
    swap/.append code={
      \ifpgfarrowswap
        \pgfkeysalso{anchor=west}
      \else
        \pgfkeysalso{anchor=east}
      \fi}},
}}
\tikzset{graphs/graph style/.style={
  tree layout,
  level sep=1cm,
  nodes={draw, rounded corners=1mm},
  edges={edge style},
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_{0}$}
}}

@@

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
\autoref{sub:Single-User Actions} explores edits by a single user, Alice,
and \autoref{sub:Multi-user Edits} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

TODO: labels for edge numbers

@@

TODO: move subcaptions into master caption

@@

TODO: Remove subfigure captions

@@

Program state is a directed graph with a distinguished root vertex.
%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)}, respectively.
The root vertex has the distinguished identifier~0 and the constructor~\textbullet.

Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{left} and~\texttt{right} children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
for the purposes of this paper, we represent identifiers and number literals
indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
The root vertex constructor,~\textbullet, has the single child position \texttt{root}.

Each edge is labeled with a globally unique identifier (e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
a child position (e.g., \texttt{left} and \texttt{root} in \autoref{fig:A simple tree with a hole-b}).
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
%Visually, we indicate the position of an edge by the location of its origin.
For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

@@

In the remainder of this section,
we will consider user actions, starting
in this subsection with single-user
actions and continuing in \autoref{sub:Multi-user Edits}
with multi-user actions.

Each user action that we consider corresponds to one or more graph edits.
Graph edits can only add fresh edges and vertices or remove existing edges.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{minipage}[t]{0.49\columnwidth}
\centering
\newedge{ConstructionY}{R}%
\newvertex{ConstructionY}{y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   "\vConstructionY" [> "\eConstructionY"]
  }
 }
};
\end{tikzpicture}
\caption{Construction.
\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{\vidConstructionY} and \Edge{\eidConstructionY}.}
\label{fig:Construction}
\label{fig:Construction-b}
\end{minipage}
%%%%%%%%
\hfill
%%%%%%%%
\begin{minipage}[t]{0.49\linewidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   {},
   "\vConstructionY" [> "\eConstructionY"]
  }
 }
};
\path (-1cm,-2.5cm) graph[graph style] {
 "\vSimpleX" % [as={$\texttt{y}_{6}$}]
};
\end{tikzpicture}
\caption{Change.}
\label{fig:Change-a}
\label{fig:Change}
\end{minipage}
%%%%%%%%
\hfill{}
\end{figure}

\todo{"user actions" and "graph edits"}
To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole}
and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
This action corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly
created vertex, \Vertex{\vidConstructionY}, containing the variable reference \texttt{y}.
The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.

@@

\todo{Arrows between figures}
Next Alice decides to delete~\texttt{x} so that the code becomes~\texttt{\hole  * y}.
This is modeled by deleting \Edge{\eidConstructionY} as shown in \autoref{fig:Change-a}. 
Notice that \Vertex{\vidConstructionY} continues to exist, and if it had any children, those children could remain connected to it.
In our system, once a vertex is created it is never deleted.
This allows further manipulations of those vertices by other users as shown later in this section.
Note that we omit such orphans from the remaining diagrams if they are not relevant to the exposition.

Once an edge with a particular identifier is deleted, it cannot be recreated.
Thus if Alice performed an "undo" on this deletion, Grove would create a fresh edge between \Vertex{TODO} and \Vertex{TODO}.

@@

\begin{figure}
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (-0.5,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-b}
\newedge{WrapPlus}{Root}
\newvertex{WrapPlus}{+}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (-0.5cm,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-c}
\newedge{WrapTimes}{L}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eWrapTimes"'] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vChangeZ" [> "\eChangeZ"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping.
\autoref{fig:Wrapping-a}:~\autoref{fig:Change-b} after deleting \Edge{\eidSimpleTimes}.
\autoref{fig:Wrapping-b}:~\autoref{fig:Wrapping-a} after adding \Vertex{\vidWrapPlus} and \Edge{\eidWrapPlus}.
\autoref{fig:Wrapping-c}:~\autoref{fig:Wrapping-b} after adding \Edge{\eidWrapTimes}.}
\label{fig:Wrapping}
\end{figure}

@@

TODO: revise: "user action" "graph edit"

Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
If this were a hole, this would cause a construction edit to be performed as in Sec.TODO above.
However, this position is already occupied by the \texttt{*} expression, 
so instead, typing~\texttt{+} performs a wrapping edit.
This corresponds to the following sequence of graph actions, shown in \autoref{fig:Wrapping}: 
(a) delete \Edge{\eidSimpleTimes}, leaving \Vertex{\vidSimpleTimes} temporarily orphaned,
(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex,
(c) add an edge to \Vertex{\vidSimpleTimes} in the left child position of this new~\texttt{+} vertex.
(The choice of wrapping with a bias for the left child position is arbitrary.)

% A naive editor could add the~\texttt{+} as a new child of the root.
% However,

by creating a new
edge from the \texttt{L} index of the \texttt{*} to a freshly created \texttt{plus} vertex.
This would result in a conflict between the \texttt{+} and the existing \texttt{var}
as shown in Figure~REF:TODO.

You can have a second edge,
we discuss conflicts in section TODO
In the single-user setting, we don't want to create conflicts, so instead we perform a wrapping action.

How to handle multiple children in the same position is explored in Sec~REF:TODO,
but for this sort of edit, a better user experience
would be interpret the request for a plus to mean to
take whatever (if anything) is in the cursor
and place it as a child of the plus.
This \emph{wraps} the existing \texttt{var} in the \texttt{plus}
and results in the code~\texttt{(x * z) + \hole}.
(This is similar to the behavior in other structure editors such as in CITE:TODO.)
Translating this into edge deletes and adds results in the sequence of graphs in \autoref{fig:Wrapping}.

TODO: "Note that the logic of to insert or wrap is a layer above graph actions."

@@

\begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:Move-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (0cm,-3cm) graph [graph style] {
 "\vSimpleTimes" -> {
  "\vSimpleX" [> "\eSimpleX"'],
  "\vChangeZ" [> "\eChangeZ"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.45\linewidth}
\centering
\caption{}
\label{fig:Move-b}
\newedge{MoveTimes}{R}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vChangeZ" [> "\eChangeZ"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.
\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{\eidWrapTimes}.
\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{\eidMoveTimes}.}
\label{fig:Move}
\end{figure}

@@

For example, consider \autoref{fig:Move}, which represents
the movement of~\texttt{1 * 2} from the left child of \texttt{plus}
to its right child.
\autoref{fig:Editing Nested Parts of the Code-c} is the code before the move.
First, \Edge{\eidWrapPlus} is deleted, which results in \autoref{fig:Move-a}
where the rooted code is now~\texttt{\hole{} + \hole{}},
and \Vertex{\vidSimpleTimes} is orphaned.
Next \Edge{\eidMoveTimes} is added, which results in the bottom graph
where the rooted code is now~\texttt{\hole{} + 1 * 2},
and \Vertex{\vidSimpleTimes} is no longer orphaned.

@@

\subsection{Multi-user Edits}
\label{sub:Multi-user Edits}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-a}
\newedge{DifferentAlice}{L}
\newvertex{DifferentAlice}{u}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vChangeZ" [> "\eChangeZ"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-b}
\newedge{DifferentBob}{R}
\newvertex{DifferentBob}{v}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Different Parts of the Code.
\textbf{(a)}:~\autoref{fig:Move-b} after Alice replaces \Vertex{\vidSimpleX} with \Vertex{\vidDifferentAlice}.
\textbf{(b)}:~\autoref{fig:Move-b} after Bob replaces \Vertex{\vidConstructionY} with \Vertex{\vidDifferentBob}.
\textbf{(c)}:~The
result of merging \autoref{fig:Editing Different Parts of the Code-a}
with \autoref{fig:Editing Different Parts of the Code-b}.}
\label{fig:Editing Different Parts of the Code}
\end{figure}

>>>>

\tikzset{
  %every picture/.style={framed, background rectangle/.style={draw=gray!50}},
}
\tikzset{edge style/.style={
  draw,
  font={\small\ttfamily},
  /tikz/every edge quotes/.style={
    %draw=gray!20,
    anchor=west,
    swap/.append code={
      \ifpgfarrowswap
        \pgfkeysalso{anchor=west}
      \else
        \pgfkeysalso{anchor=east}
      \fi}},
}}
\tikzset{graphs/graph style/.style={
  tree layout,
  level sep=1cm,
  nodes={draw, rounded corners=1mm},
  edges={edge style},
}}
\tikzset{graphs/root style/.style={
 %draw=none,
 as={\textbullet$_{0}$}
}}
\tikzset{alice/.style={
  color=red,
}}

@@

This section introduces collaborative structure editing in Grove.
We begin in \autoref{sub:Program Representation} by covering how we use graphs to represent expressions with holes,
\autoref{sub:Single-User Actions} explores edits by a single user, Alice,
and \autoref{sub:Multi-user Actions} explores edits by multiple users, Alice and Bob,
each of whom is editing their own instance of a program.

@@

@@

@@

@@

Program state is a directed graph with a distinguished root vertex.
%For example, \autoref{fig:A simple tree with a hole-b} shows the graph corresponding to the term in \autoref{fig:A simple tree with a hole-a}.
Each vertex represents a term and is labeled with both a globally unique identifier and a term constructor.
In \autoref{fig:A simple tree with a hole-b}, \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX} have the constructors~\texttt{*} and \texttt{var(x)},\todo{var} respectively.
The root vertex has the distinguished identifier~0 and the constructor~\textbullet.

Each term constructor has an associated set of child positions.
The~\texttt{*} constructor has positions for~\texttt{L}~(i.e., left) and~\texttt{R}~(i.e., right) children
and the~\texttt{var} constructors have no child positions but takes an argument naming the particular
variable referenced~(e.g.,~\texttt{x} in \texttt{var(x)}).\footnote{Note that,
for the purposes of this paper, we represent identifiers and number literals
indivisibly.  See Section~TODO:REF, for how we would allow editing individual characters.}
The root vertex constructor,~\textbullet, has the single child position \texttt{Root}.

Each edge is labeled with a globally unique identifier~(e.g.,~1 and~3 in \autoref{fig:A simple tree with a hole-b}) and
a child position (e.g., \texttt{L} and \texttt{Root} in \autoref{fig:A simple tree with a hole-b}).
An edge indicates that the destination vertex is a child of the origin vertex at the given position.
%Visually, we indicate the position of an edge by the location of its origin.
For the sake of presentation clarity, we reserve odd identifiers for vertexes and even identifiers for edges.

@@

In the remainder of this section,
we consider user actions, starting
in this subsection with single-user
actions and continuing in \autoref{sub:Multi-user Actions}
with multi-user actions.

Each user action corresponds to one or more graph edits.
Graph edits can only add fresh edges and vertices or remove existing edges.

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{minipage}[t]{0.49\columnwidth}
\centering
\newedge{ConstructionY}{R}%
\newvertex{ConstructionY}{y}%
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   "\vSimpleX" [> "\eSimpleX"'],
   "\vConstructionY" [> "\eConstructionY", >alice, alice]
  }
 }
};
\end{tikzpicture}
\caption{Construction.
\autoref{fig:Construction-b}:~\autoref{fig:A simple tree with a hole-b} after adding \Vertex{\vidConstructionY} and \Edge{\eidConstructionY}.}
\label{fig:Construction}
\label{fig:Construction-b}
\end{minipage}
%%%%%%%%
\hfill
%%%%%%%%
\begin{minipage}[t]{0.49\linewidth}
\centering
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vSimpleTimes" [> "\eSimpleTimes"] -> {
   {},
   "\vConstructionY" [> "\eConstructionY"]
  }
 }
};
\path (-0.75cm,-2.75cm) graph[graph style] {
 "\vSimpleX" % [as={$\texttt{y}_{6}$}]
};
\end{tikzpicture}
\caption{Change.}
\label{fig:Change-a}
\label{fig:Change}
\end{minipage}
%%%%%%%%
\hfill{}
\end{figure}

\todo{"user actions" and "graph edits"}
To start our examples, Alice moves her cursor to the hole in~\texttt{x * \hole} in \autoref{fig:A simple tree with a hole}
and constructs the variable~\texttt{y} as shown in \autoref{fig:Construction-b}.
This action corresponds to creating \Edge{\eidConstructionY} from \Vertex{\vidSimpleX} at the \texttt{R} position to a newly
created vertex, \Vertex{\vidConstructionY}, containing the variable reference~\texttt{y}.
The resulting graph, shown in \autoref{fig:Construction-b}, represents the expression \texttt{x * y}.

@@

\todo{Arrows between figures}
Next Alice deletes~\texttt{x} so that the code becomes~\texttt{\hole{} * y}.
This is modeled by deleting \Edge{\eidConstructionY} as shown in \autoref{fig:Change-a}. 
Notice that \Vertex{\vidConstructionY} continues to exist, and if it had any children, those children would remain connected to it.
In our system, once a vertex is created it is never deleted.
This allows further manipulations of those vertices by other users as shown later in this section.
Note that we omit such orphaned vertexes from the remaining diagrams if they are not relevant to the exposition.

Once an edge with a particular identifier is deleted, it cannot be recreated.
Thus if Alice performed an ``undo'' on this deletion, Grove would create a fresh edge between \Vertex{\vidSimpleTimes} and \Vertex{\vidSimpleX}.

@@

\begin{figure}
% TODO: naive add that creates a conflict
% \begin{tikzpicture}
% \path (0,0)       node {\includegraphics[scale=0.4]{graphviz/wrapping/1.dot}};
% \path (0cm,-4cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/2.dot}};
% \path (0cm,-8cm)  node {\includegraphics[scale=0.4]{graphviz/wrapping/3.dot}};
% \path (0cm,-12cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/4.dot}};
% \path (0cm,-16cm) node {\includegraphics[scale=0.4]{graphviz/wrapping/5.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style]
};
\path (-0.5cm,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  {},
  "\vConstructionY" [> "\eConstructionY"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-b}
\newedge{WrapPlus}{Root}
\newvertex{WrapPlus}{+}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus", >alice, alice]
 }
};
\path (-0.5cm,-3cm) graph[graph style] {
 "\vSimpleTimes" -> {
  {},
  "\vConstructionY" [> "\eConstructionY"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Wrapping-c}
\newedge{WrapTimes}{L}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   "\vSimpleTimes" [> "\eWrapTimes"', >alice] -> {
    {},
    "\vConstructionY" [> "\eConstructionY"]
   },
   {}
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Wrapping.}
\label{fig:Wrapping}
\end{figure}

@@

Next, Alice types~\texttt{+} when the cursor is on \Vertex{\vidSimpleTimes}.
If this were a hole, this would cause a construction edit to be performed as in \autoref{sub:Construction} above.
However, this position is already occupied by the \texttt{*} expression, 
so instead, typing~\texttt{+} is interpreted as a wrapping action.\footnote{%
Multi-user merges can still create multi-child conflicts,
but it is better to not interpret user actions ina way that creates them.
How to handle multiple children in the same position is explored in \autoref{sub:Multi-child conflicts}.}
This corresponds to the following sequence of graph edits, shown in \autoref{fig:Wrapping}: 
(a) delete \Edge{\eidSimpleTimes}, leaving \Vertex{\vidSimpleTimes} temporarily orphaned,
(b) add a fresh~\texttt{+} vertex (\Vertex{\vidWrapPlus}) as a child of the root vertex, and
(c) add an edge to \Vertex{\vidSimpleTimes} in the left child position of this new~\texttt{+} vertex.
(The choice of wrapping with a bias for the left child position is arbitrary.)

@@

\begin{figure}
% \centering
% \begin{tikzpicture}
% \path (0,0)    node {\includegraphics[scale=0.75]{graphviz/move/1.dot}};
% \path (0,-5cm) node {\includegraphics[scale=0.75]{graphviz/move/2.dot}};
% \path (0,-9cm) node {\includegraphics[scale=0.75]{graphviz/move/3.dot}};
% \end{tikzpicture}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.49\linewidth}
\centering
\caption{}
\label{fig:Move-a}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"]
 }
};
\path (0cm,-3cm) graph [graph style] {
 "\vSimpleTimes" -> {
  {},
  "\vConstructionY" [> "\eConstructionY"]
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.45\linewidth}
\centering
\caption{}
\label{fig:Move-b}
\newedge{MoveTimes}{R}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    {},
    "\vConstructionY" [> "\eConstructionY"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Moving \Vertex{\vidSimpleTimes} from the left child to the right child of \Vertex{\vidWrapPlus}.
\autoref{fig:Move-a}:~\autoref{fig:Wrapping-c} after deleting \Edge{\eidWrapTimes}.
\autoref{fig:Move-b}:~\autoref{fig:Move-a} after adding \Edge{\eidMoveTimes}.}
\label{fig:Move}
\end{figure}

@@

For example, \autoref{fig:Move} Alice
moves \texttt{\hole{} * y} from the left child of \texttt{+} to its right child.
This corresponds to the following sequence of graph edits, shown in \autoref{fig:Move}:
(a) delete \Edge{\eidWrapPlus} leaving \Vertex{\vidSimpleTimes} temporarily orphaned and 
(b) add an edge (\Edge{\eidMoveTimes}) to \Vertex{\vidSimpleTimes} in the right child position of \Vertex{\vidWrapPlus}.

@@

\subsection{Multi-user Actions}
\label{sub:Multi-user Actions}

@@

\begin{figure}
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-a}
\newedge{DifferentAlice}{L}
\newvertex{DifferentAlice}{u}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vConstructionY" [> "\eConstructionY"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-b}
\newedge{DifferentBob}{R}
\newvertex{DifferentBob}{v}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vSimpleX" [> "\eSimpleX"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill
%%%%%%%%
\begin{subfigure}[t]{0.32\linewidth}
\centering
\caption{}
\label{fig:Editing Different Parts of the Code-c}
\begin{tikzpicture}
\path (0cm,0cm) graph[graph style] {
 root[root style] -> {
  "\vWrapPlus" [> "\eWrapPlus"] -> {
   {},
   "\vSimpleTimes" [> "\eMoveTimes"] -> {
    "\vDifferentAlice" [> "\eDifferentAlice"'],
    "\vDifferentBob" [> "\eDifferentBob"]
   }
  }
 }
};
\end{tikzpicture}
\end{subfigure}
%%%%%%%%
\hfill{}
\caption{Editing Different Parts of the Code.
\textbf{(a)}:~\autoref{fig:Move-b} after Alice replaces \Vertex{\vidSimpleX} with \Vertex{\vidDifferentAlice}.
\textbf{(b)}:~\autoref{fig:Move-b} after Bob replaces \Vertex{\vidConstructionY} with \Vertex{\vidDifferentBob}.
\textbf{(c)}:~The
result of merging \autoref{fig:Editing Different Parts of the Code-a}
with \autoref{fig:Editing Different Parts of the Code-b}.}
\label{fig:Editing Different Parts of the Code}
\end{figure}
